---
title: Random Seed Group Lasso Cross-Validation
author: Lizzy Gibson
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(grpreg)
library(janitor)
library(glmnet)

#This turns off scientific notation
options(scipen = 999)

knitr::opts_chunk$set(
	echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Data Import 

First, load the dataset; clean up names as needed; and convert factors to, well, factors. 

```{r}
study_pop = read_csv("Data/studypop.csv") %>% 
  clean_names(case = c("old_janitor")) %>% 
  mutate(bmi_cat3 = as.factor(bmi_cat3),
         edu_cat = as.factor(edu_cat),
         race_cat = as.factor(race_cat),
         male = as.factor(male)) 

data_lasso = study_pop %>% 
  mutate_at(vars(contains("la")), log) %>% 
  mutate(log_telomean = log(telomean)) %>% 
  dplyr::select(log_telomean, lbx074la:lbx187la, lbxd03la:lbx194la, everything(), -seqn, -telomean) %>% 
  na.omit(log_telomean)

data_lasso

# Create a matrix of predictors as x
x = model.matrix(log_telomean ~ ., data_lasso)[,-1]

# Extract outcome vector
y = data_lasso$log_telomean
```

## Group Definitions

Create grouping variable. If there are coefficients to be included in the model without being penalized, assign them to group 0 (or "0"). This includes all covariates we want to keep in the model. Group must be a FACTOR. 

* 3 groups = 
    * 8 non-Dioxin-like PCBs
    * 2 non-ortho PCBs
    * TEQ (3 dioxins, 4 furans, 1 mono-ortho (Dioxin-like) PCBs). 

```{r}
group3 <- vector()
group3[grepl("lbx1|0",colnames(x))] <- "Non-Dioxin-like PCB"
group3[grepl("lbxd", colnames(x))] <- "TEQ"
group3[grepl("lbxf", colnames(x))] <- "TEQ"
group3[grepl("lbxh|p", colnames(x))] <- "Non-Ortho PCB"
group3[grepl("lbx118la", colnames(x))] <- "TEQ"
group3[grepl("pct", colnames(x))] <- "0"
group3[grepl("bcsi", colnames(x))] <- "0"
group3[grepl("bmi", colnames(x))] <- "0"
group3[grepl("edu", colnames(x))] <- "0"
group3[grepl("race", colnames(x))] <- "0"
group3[grepl("male", colnames(x))] <- "0"
group3[grepl("bxcot", colnames(x))] <- "0"
group3[grepl("age", colnames(x))] <- "0"
group3 <- as.factor(group3)

cbind(colnames(x), group3) #bind by columns
```

## Grouped Variable Selection

```{r models_3}
lasso_3 <- grpreg(x, y, group3, penalty = "grLasso")
```

Results are illustrated below. 

```{r}
plot(lasso_3)
```

## Lasso w/ CV

As with `glmnet`, we can use CV to choose the tuning parameters. If not specified in code, ten-fold CV is the default. 

We perform this analysis with the three-group model and the lasso penalty.

```{r cvlasso}
cv_lasso_3 <- cv.grpreg(x, y, group3, 
                        penalty = "grLasso", seed = 1988)

plot(cv_lasso_3)

cv_lasso_3$cve #cross-validation error
cv_lasso_3$cvse #std error of CV error
log(cv_lasso_3$lambda) # log(lambda)

plot(desc(log(cv_lasso_3$lambda)), cv_lasso_3$cve)
```

```{r}
#Find the lambda that results in the smallest CV error
best_lasso_3 <- cv_lasso_3$lambda.min

#Fit model with cross-validated lambda
fit_lasso_3 <- grpreg(x, y, group3, penalty = "grLasso", lambda = best_lasso_3)

#Look at coefficients from model with best lambda
fit_lasso_3$beta

##prediction and MSE using train dataset
grlasso.pred <-  predict(fit_lasso_3, x)
grlasso_MSE <-  mean((grlasso.pred - y)^2)
grlasso_MSE
```

### Data Visualization

```{r vizlasso}
#3 group model
lasso_beta_3 <- as_tibble(fit_lasso_3$beta, rownames = "variable") %>% 
  rename(beta = 2) %>% 
  filter(variable != "(Intercept)") %>% 
  mutate(group3 = group3,
         method = "Grouped Lasso")

#Filter out covariates
lasso_beta_3 %>% 
  filter(group3 %in% c("TEQ", "Non-Dioxin-like PCB", "Non-Ortho PCB")) %>% 
  ggplot(aes(x = variable, y = round(beta, 5))) + 
  geom_point() + theme_minimal() + geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  facet_wrap(~ group3, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() + 
  labs(title = "Lasso Coefficients for 3 Group Model", y = "Beta Coefficient", x = "Variable")
```

## Loop Over Seeds

```{r}
fit_with_cv = function(seed) {
  cv_lasso_loop <- cv.grpreg(x, y, group3, 
                          penalty = "grLasso", seed = seed)
  
  loop_cve   <- cv_lasso_loop$cve  #cross-validation error
  loop_cvse  <- cv_lasso_loop$cvse #std error of CV error
  loop_lam <- cv_lasso_loop$lambda 
  
  dat <- cbind(lam = loop_lam, cve = loop_cve, cvse = loop_cvse) %>% 
    as_tibble() %>% 
    mutate(upper = cve + 1.96*cvse,
         lower = cve - 1.96*cvse) %>% 
    mutate(seed = seed)
}

repeat_cv = 
  tibble(seed = 1:5) %>%
  mutate(
    fits = map(seed, ~fit_with_cv(seed = .x))) %>% 
  unnest()
  
repeat_cv %>% 
  ggplot(aes(x = lam, y = cve)) + 
  geom_pointrange(aes(group = seed, ymin = lower, ymax = upper)) + 
  geom_line(aes(group = seed))

```

