---
title: "beta coefficients plot"
author: "Yanelli Nunez"
date: "7/12/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


-If I give you seed 3 I get a full range and if I give you seed 1 I get a single point. 
-Do group lasso with single element groupings and see if we can replicate lasso
-what is the vector matrix structure that we suppose to use for grp lasso 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(caret)
library(Hmisc)
library(glmnet)
library(grpreg)
```

Note: R is version: 3.6.1
glmnet and grpreg package are the new versions
```{r}
sessionInfo()
```

### Data Import and Cleaning 

First, load the dataset; clean up names as needed; and convert factors to, well, factors. Next we remove missing values and reorder predictors (environmental variables first, confounders second). In keeping with standard practice, we'll ln-transform the environmental exposures and the outcome. This is the dataset we'll use to illustrate variable selection methods.

```{r}
study_pop = read_csv("./Data/studypop.csv") %>% 
  clean_names(case = c("old_janitor")) %>% 
  mutate(bmi_cat3 = as.factor(bmi_cat3),
         edu_cat = as.factor(edu_cat),
         race_cat = as.factor(race_cat),
         male = as.factor(male)) 

data = study_pop %>% 
  mutate_at(vars(contains("la")), log) %>% 
  mutate(log_telomean = log(telomean)) %>% 
  dplyr::select(log_telomean, lbx074la:lbx187la, lbxd03la:lbx194la, everything(), -seqn, -telomean) %>% 
  na.omit(log_telomean) 

# Create a matrix of predictors as x
x = model.matrix(log_telomean ~ ., data)[,-1]

# Extract outcome vector
y = data$log_telomean
```

# Lasso 
### Lasso w/ CV

Some built-in functions will conduct a cross-validation analysis and identify the "best" tuning parameter.
Optional arguments to `glmnet` can be useful -- in particular, `weights` can be used in the context of the adaptive lasso and `penalty.factor` can separate penalized variables from confounders. 

```{r}
set.seed(2)

is_penalized = c(rep(1, ncol(x[,1:18])), rep(0, ncol(x[,19:36])))

# Use cross-validation (CV) to find best lambda value
cv.lasso = cv.glmnet(x, y, 
                     penalty.factor = is_penalized,
                     type.measure = "mse", alpha = 1)
plot(cv.lasso)

best_lambda = cv.lasso$lambda.min
best_lambda
```


### Looping seeds for lasso for coefficients plot
```{r}
fit_cv_lasso = function(seed) {
  set.seed(seed)
  cv.lasso = cv.glmnet(x, y, 
                     penalty.factor = c(rep(1, ncol(x[,1:18])), rep(0, ncol(x[,19:36]))),
                     type.measure = "mse", alpha = 1)
  
  #Find the lambda that results in the smallest CV error
  best_lambda <- cv.lasso$lambda.min
 
  
  #Fit model with cross-validated lambda
  lasso.mod = glmnet(x, y, 
                   penalty.factor = c(rep(1, ncol(x[,1:18])), rep(0, ncol(x[,19:36]))),
                   alpha = 1, lambda = best_lambda)

lasso.mod
}


extract_coefs = function(fit) {
 as.tibble(as.matrix(fit$beta), rownames = "variable") %>%
    rename(beta = s0) 
}


repeat_cv_lasso = 
  tibble(seed = 1:100) %>% 
  mutate(
    fits = map(seed, ~fit_cv_lasso(seed = .x)),
    coefs = map(fits, extract_coefs)) %>% 
  dplyr::select(-fits) %>% 
  unnest() %>%
  rename(beta_lasso = "beta") 


```

### Looping seeds for lasso for lambda plot
```{r}

fit_cv_lasso_2 = function(seed) {
  set.seed(seed)
  cv.lasso = cv.glmnet(x, y, 
                     penalty.factor = c(rep(1, ncol(x[,1:18])), rep(0, ncol(x[,19:36]))),
                     type.measure = "mse", alpha = 1)

  
  loop_cvm    <- cv.lasso$cvm  # mean cross-validation error
  loop_cvsd   <- cv.lasso$cvsd #std error of CV error
  loop_lam    <- cv.lasso$lambda 
  best_lambda <- cv.lasso$lambda.min # to plot min, median, max
  
  dat <- cbind(lam = loop_lam, cvm = loop_cvm, cvsd = loop_cvsd, best_lambda = best_lambda) %>% 
    as_tibble() %>% 
    mutate(upper = cvm + 1.96*cvsd,
         lower = cvm - 1.96*cvsd) %>% 
    mutate(seed1 = seed)

}

repeat_cv_lasso_2 = 
  tibble(seed = 1:100) %>%
  mutate(
    fits = map(seed, ~fit_cv_lasso_2(seed = .x))) %>% 
  unnest() %>% 
  mutate(seed = as.factor(seed))

  
```


# Lasso visualization

#### betas
```{r}

repeat_cv_lasso  %>%
  ggplot() + 
  geom_point(aes(x = seed, y = beta_lasso), color = 'red', size = 0.5, alpha = 0.5) + 
  facet_wrap(~variable) + 
  geom_hline(yintercept = 0) +
  ggtitle("R3.6.1 glmnet 2.0-18 betas")

```

#compiling data for final summary beta plot
```{r}
data_lasso <- repeat_cv_lasso %>%
  filter(!variable %in% c("bmi_cat32", "bmi_cat33", "edu_cat2", "edu_cat3", "edu_cat4", "race_cat2", "race_cat3", "race_cat4", "male1", "ln_lbxcot", "age_cent", "age_sq", "lbxwbcsi", "lbxlypct", "lbxlypct", "lbxmopct", "lbxmopct", "lbxnepct",
"lbxnepct", "lbxeopct", "lbxeopct", "lbxbapct", "lbxbapct")) %>%
  mutate(Group = ifelse(variable == "lbx118la", "mPFD", 
                        ifelse(grepl("lbx1", variable), "Non-Dioxin-like PCBs",
                          ifelse(grepl("lbx0", variable), "Non-Dioxin-like PCBs",
                            ifelse(grepl("lbxp", variable), "Non-Ortho PCBs",
                              ifelse(grepl("lbxh", variable), "Non-Ortho PCBs", "mPFD")))))) %>%
  mutate(variable = as.factor(variable),
          chem = fct_recode(variable, 
                             "1,2,3,6,7,8-hxcdd" = "lbxd03la",
                                "1,2,3,4,6,7,8-hpcdd" = "lbxd05la",
                               "1,2,3,4,6,7,8,9-ocdd" =  "lbxd07la",
                               "2,3,4,7,8-pncdf" =  "lbxf03la",
                               "1,2,3,4,7,8-hxcdf" =  "lbxf04la",
                               "1,2,3,6,7,8-hxcdf" =  "lbxf05la",
                               "1,2,3,4,6,7,8-hxcdf" =  "lbxf08la",
                                "PCB 74" = "lbx074la",
                                "PCB 99" = "lbx099la",
                                "PCB 118" = "lbx118la",
                                "PCB 138" = "lbx138la",
                                "PCB 153" = "lbx153la",
                                "PCB 170" = "lbx170la",
                                "PCB 180" = "lbx180la",
                                "PCB 187" = "lbx187la",
                                "PCB 194" = "lbx194la",
                                "PCB 126" = "lbxpcbla",
                                "PCB 169" =  "lbxhxcla")) %>%
  dplyr::select(-variable) %>%
  group_by(chem) %>%
  mutate(mean = mean(beta_lasso),
            se = sd(beta_lasso)/sqrt(length(beta_lasso))) %>%
  dplyr::select(seed, Group, chem, beta_lasso, everything())

#counting non zero coefficients

counts <- data_lasso %>%
  filter(beta_lasso != 0) %>%
  group_by(chem) %>%
  summarise(number_nonzero = n())
  

summary_data_lasso <- full_join(data_lasso, counts, by = "chem") %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))


summary_data_lasso$chem <- factor(summary_data_lasso$chem,levels = c("1,2,3,6,7,8-hxcdd",
                                "1,2,3,4,6,7,8-hpcdd",
                               "1,2,3,4,6,7,8,9-ocdd",
                               "2,3,4,7,8-pncdf",
                               "1,2,3,4,7,8-hxcdf",
                               "1,2,3,6,7,8-hxcdf",
                               "1,2,3,4,6,7,8-hxcdf",
                               "PCB 118",
                               "PCB 74",
                                "PCB 99",
                                "PCB 138",
                                "PCB 153",
                                "PCB 170",
                                "PCB 180",
                                "PCB 187",
                                "PCB 194",
                                "PCB 126",
                                "PCB 169"))

#write.csv(summary_data_lasso, "summary_data_lasso.csv")


```

#final summary plot for lasso
```{r}
summary_data_lasso <- read_csv("./lasso_glmnet/summary_data_lasso.csv") %>%
  dplyr::select(-X1)

summary_data_lasso$chem <- factor(summary_data_lasso$chem,levels = c("1,2,3,6,7,8-hxcdd",
                                "1,2,3,4,6,7,8-hpcdd",
                               "1,2,3,4,6,7,8,9-ocdd",
                               "2,3,4,7,8-pncdf",
                               "1,2,3,4,7,8-hxcdf",
                               "1,2,3,6,7,8-hxcdf",
                               "1,2,3,4,6,7,8-hxcdf",
                               "PCB 118",
                               "PCB 74",
                                "PCB 99",
                                "PCB 138",
                                "PCB 153",
                                "PCB 170",
                                "PCB 180",
                                "PCB 187",
                                "PCB 194",
                                "PCB 126",
                                "PCB 169"))


summary_data_lasso %>%
  group_by(chem) %>%
  ggplot(aes(x = chem)) +
   geom_hline(yintercept = 0, color = "black") +
  geom_jitter(aes(y = beta_lasso, color = Group), shape=16, size = 1, alpha = 1,  position=position_jitter(width = .2)) +
  #geom_point(aes(y = mean, fill = Group), shape = 23, size = 3, alpha = 1) +
 #geom_errorbar(aes(ymin = mean-se, ymax=mean+se), width=.1, position=position_dodge(0.05)) +
 #stat_summary(aes(y = beta_lasso), fun.data = mean_sdl, fun.args = list(mult = 1), geom = "errorbar", width=0.25) +
  # stat_summary(aes(y = beta_lasso, fill = Group), fun.y = mean, geom ="point", shape=23, size=3, alpha = 1) +
  geom_col(aes(y = number_nonzero/4000, fill = Group), width = 0.8, position ="identity", alpha = 0.002) +
  scale_y_continuous(sec.axis = sec_axis(~.*4000, name = "# of Non-zero Beta Coefficients")) +
  theme_classic() +
  labs(y = "Beta Coefficients",
       x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 9.5),
                      legend.position = "bottom",
                      legend.text = element_text(size = 12),
        axis.text.y = element_text(size = 9.5),
        text = element_text(size = 12.5), 
        aspect.ratio = 1/1,
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_line(color = "black")) +
  geom_boxplot(aes(y = beta_lasso), outlier.shape = NA, alpha = 0) 
 
  
```


### Lambda with Error Bars
```{r}
repeat_cv_lasso_2 %>% 
  ggplot(aes(x = lam, y = cvm, color = seed)) + 
  geom_point() + 
  geom_line() +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  theme(legend.position = "none") +
  labs(x = expression(lambda), y = "Cross-Validation Error")
```


### Lambda without Error Bars
```{r}
repeat_cv_lasso_2 %>% 
  ggplot(aes(x = lam, y = cvm, color = seed)) + 
  geom_point() + 
  geom_line() +
  theme(legend.position = "none") +
  labs(x = expression(lambda), y = "Cross-Validation Error") +
  ggtitle("R3.6.1 glmnet 2.0-18 CV")
```

#### Lambda to Plot
```{r}
repeat_cv_lasso_2 %>% dplyr::select(seed, best_lambda) %>% distinct() %>% summary()
repeat_cv_lasso_2 %>% dplyr::select(seed, best_lambda) %>% distinct() %>% pull(best_lambda) %>% quantile()

pdf("./Figures/cv_lasso.pdf")
repeat_cv_lasso_2 %>% 
  ggplot(aes(x = lam, y = cvm)) + 
  geom_line(aes(group = seed), color = "gray70") +
  geom_rect(aes(xmin = 0.0008442, xmax = 0.0037404, 
               ymin = -Inf, ymax = Inf),
               fill = "pink", 
               alpha = 0.01) +
  geom_vline(xintercept = 0.0031053, color = "red") + # Median
  geom_vline(xintercept = 0.0008442, color = "red", linetype = "longdash") + # Min
  geom_vline(xintercept = 0.0037404, color = "red", linetype = "longdash") + # Max
  geom_vline(xintercept = 0.0028294720, color = "red", linetype = "dashed") + # Q25
  geom_vline(xintercept = 0.0034081090, color = "red", linetype = "dashed") + # Q75
  theme_classic(base_size = 15) +
  theme(legend.position = "none") +
  labs(x = expression(lambda), y = "Cross-Validation Error")
dev.off()
```


### Selected lambda
```{r}
min_cv <- repeat_cv_lasso_2 %>% 
  group_by(seed) %>% 
  filter(cvm == min(cvm))

min_cv %>% 
  ggplot(aes(y = lam, x = seed)) + 
  geom_point(aes(color = lam)) + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, size = 7)) +
  labs(x = "seed", y = expression("Cross-Validation Selected "*lambda))
```


