---
title: "beta coefficients plot"
author: "Yanelli Nunez"
date: "7/12/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---


-If I give you seed 3 I get a full range and if I give you seed 1 I get a single point. 
-Do group lasso with single element groupings and see if we can replicate lasso
-what is the vector matrix structure that we suppose to use for grp lasso 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(caret)
library(Hmisc)
library(glmnet)
library(grpreg)
```

Note: R is version: 3.6.1
glmnet and grpreg package are the new versions
```{r}
sessionInfo()
```

### Data Import and Cleaning 

First, load the dataset; clean up names as needed; and convert factors to, well, factors. Next we remove missing values and reorder predictors (environmental variables first, confounders second). In keeping with standard practice, we'll ln-transform the environmental exposures and the outcome. This is the dataset we'll use to illustrate variable selection methods.

```{r}
study_pop = read_csv("Data/studypop.csv") %>% 
  clean_names(case = c("old_janitor")) %>% 
  mutate(bmi_cat3 = as.factor(bmi_cat3),
         edu_cat = as.factor(edu_cat),
         race_cat = as.factor(race_cat),
         male = as.factor(male)) 

data = study_pop %>% 
  mutate_at(vars(contains("la")), log) %>% 
  mutate(log_telomean = log(telomean)) %>% 
  dplyr::select(log_telomean, lbx074la:lbx187la, lbxd03la:lbx194la, everything(), -seqn, -telomean) %>% 
  na.omit(log_telomean) 

# Create a matrix of predictors as x
x = model.matrix(log_telomean ~ ., data)[,-1]

# Extract outcome vector
y = data$log_telomean
```

# Lasso 
### Lasso w/ CV

Some built-in functions will conduct a cross-validation analysis and identify the "best" tuning parameter.
Optional arguments to `glmnet` can be useful -- in particular, `weights` can be used in the context of the adaptive lasso and `penalty.factor` can separate penalized variables from confounders. 

```{r}
set.seed(2)

is_penalized = c(rep(1, ncol(x[,1:18])), rep(0, ncol(x[,19:36])))

# Use cross-validation (CV) to find best lambda value
cv.lasso = cv.glmnet(x, y, 
                     penalty.factor = is_penalized,
                     type.measure = "mse", alpha = 1)
plot(cv.lasso)

best_lambda = cv.lasso$lambda.min
best_lambda
```


### Looping seeds for lasso for coefficients plot
```{r}
fit_cv_lasso = function(seed) {
  set.seed(seed)
  cv.lasso = cv.glmnet(x, y, 
                     penalty.factor = c(rep(1, ncol(x[,1:18])), rep(0, ncol(x[,19:36]))),
                     type.measure = "mse", alpha = 1)
  
  #Find the lambda that results in the smallest CV error
  best_lambda <- cv.lasso$lambda.min
 
  
  #Fit model with cross-validated lambda
  lasso.mod = glmnet(x, y, 
                   penalty.factor = c(rep(1, ncol(x[,1:18])), rep(0, ncol(x[,19:36]))),
                   alpha = 1, lambda = best_lambda)

lasso.mod
}


extract_coefs = function(fit) {
 as.tibble(as.matrix(fit$beta), rownames = "variable") %>%
    rename(beta = s0) 
}


repeat_cv_lasso = 
  tibble(seed = 1:100) %>% 
  mutate(
    fits = map(seed, ~fit_cv_lasso(seed = .x)),
    coefs = map(fits, extract_coefs)) %>% 
  dplyr::select(-fits) %>% 
  unnest() %>%
  rename(beta_lasso = "beta") 


```

### Looping seeds for lasso for lambda plot
```{r}

fit_cv_lasso_2 = function(seed) {
  set.seed(seed)
  cv.lasso = cv.glmnet(x, y, 
                     penalty.factor = c(rep(1, ncol(x[,1:18])), rep(0, ncol(x[,19:36]))),
                     type.measure = "mse", alpha = 1)

  
 loop_cvm    <- cv.lasso$cvm  # mean cross-validation error
  loop_cvsd   <- cv.lasso$cvsd #std error of CV error
  loop_lam    <- cv.lasso$lambda 
  
  dat <- cbind(lam = loop_lam, cvm = loop_cvm, cvsd = loop_cvsd) %>% 
    as_tibble() %>% 
    mutate(upper = cvm + 1.96*cvsd,
         lower = cvm - 1.96*cvsd) %>% 
    mutate(seed1 = seed)

}

repeat_cv_lasso_2 = 
  tibble(seed = 1:100) %>%
  mutate(
    fits = map(seed, ~fit_cv_lasso_2(seed = .x))) %>% 
  unnest() %>% 
  mutate(seed = as.factor(seed))

  
```


# Lasso visualization

#### betas
```{r}

repeat_cv_lasso  %>%
  ggplot() + 
  geom_point(aes(x = seed, y = beta_lasso), color = 'red', size = 0.5, alpha = 0.5) + 
  facet_wrap(~variable) + 
  geom_hline(yintercept = 0) +
  ggtitle("R3.6.1 glmnet 2.0-18 betas")

```

#compiling data for final summary beta plot
```{r}
summary_plot <- repeat_cv_lasso %>%
  filter(!variable %in% c("bmi_cat32", "bmi_cat33", "edu_cat2", "edu_cat3", "edu_cat4", "race_cat2", "race_cat3", "race_cat4", "male1", "ln_lbxcot", "age_cent", "age_sq", "lbxwbcsi", "lbxlypct", "lbxlypct", "lbxmopct", "lbxmopct", "lbxnepct",
"lbxnepct", "lbxeopct", "lbxeopct", "lbxbapct", "lbxbapct")) %>%
  mutate(Group = ifelse(variable == "lbx118la", "mPFD", 
                        ifelse(grepl("lbx1", variable), "Non-Dioxin-like PCBs",
                          ifelse(grepl("lbx0", variable), "Non-Dioxin-like PCBs",
                            ifelse(grepl("lbxp", variable), "Non-Ortho PCBs",
                              ifelse(grepl("lbxh", variable), "Non-Ortho PCBs", "mPFD")))))) %>%
  mutate(variable = as.factor(variable),
          chem = fct_recode(variable, "PCB 74" = "lbx074la",
                                "PCB 99" = "lbx099la",
                                "PCB 118" = "lbx118la",
                                "PCB 138" = "lbx138la",
                                "PCB 153" = "lbx153la",
                                "PCB 170" = "lbx170la",
                                "PCB 180" = "lbx180la",
                                "PCB 187" = "lbx187la",
                                "PCB 194" = "lbx194la",
                                "1,2,3,6,7,8-hxcdd" = "lbxd03la",
                                "1,2,3,4,6,7,8-hpcdd" = "lbxd05la",
                               "1,2,3,4,6,7,8,9-ocdd" =  "lbxd07la",
                               "2,3,4,7,8-pncdf" =  "lbxf03la",
                               "1,2,3,4,7,8-hxcdf" =  "lbxf04la",
                               "1,2,3,6,7,8-hxcdf" =  "lbxf05la",
                               "1,2,3,4,6,7,8-hxcdf" =  "lbxf08la",
                               "PCB 169" =  "lbxhxcla",
                                "PCB 126" = "lbxpcbla")) %>%
  dplyr::select(-variable) %>%
  group_by(chem) %>%
  mutate(mean = mean(beta_lasso),
            se = sd(beta_lasso)/sqrt(length(beta_lasso))) %>%
  dplyr::select(seed, Group, chem, beta_lasso, everything())

#counting non zero coefficients

counts <- summary_plot %>%
  filter(beta_lasso != 0) %>%
  group_by(chem) %>%
  summarise(number_nonzero = n())
  

summary_data <- full_join(summary_plot, counts, by = "chem") %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))

summary(summary_data$beta_lasso)

```

#final summary beta plot
```{r}

summary_data %>%
  group_by(chem) %>%
  ggplot(aes(x = chem)) +
 geom_point(aes(y = beta_lasso, color = Group), size = 0.9, alpha = 0.8) +
  geom_point(aes(y = mean, colour = Group, fill = Group), shape = 23, size = 3, alpha = 0.3) +
 #geom_errorbar(aes(ymin = mean-se, ymax=mean+se), width=.1, position=position_dodge(0.05)) +
  stat_summary(aes(y = beta_lasso, fill = Group), fun.y = mean, geom = "point", shape=23, size=3, alpha = 0.4) +
  #geom_boxplot(aes(y = beta_lasso, color = Group), outlier.shape = NA, alpha = 0) +
  geom_col(aes(y = number_nonzero/4000, fill = Group), position ="identity", alpha = 0.005) +
  scale_y_continuous(sec.axis = sec_axis(~.*4000, name = "# of Non-zero Beta Coefficients")) +
  theme_classic() +
  labs(y = "Beta Coefficients",
       x = element_blank(),
       title  = "Lasso Beta Coefficients Over 100 seeds") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 9.5),
                      legend.position = "bottom",
                      legend.text = element_text(size = 12),
        axis.text.y = element_text(size = 9.5),
        text = element_text(size = 12.5)) +
  geom_hline(yintercept = 0) 
 
  
  
```



### Lambda with Error Bars
```{r}
repeat_cv_lasso_2 %>% 
  ggplot(aes(x = lam, y = cvm, color = seed)) + 
  geom_point() + 
  geom_line() +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  theme(legend.position = "none") +
  labs(x = expression(lambda), y = "Cross-Validation Error")
```


### Lambda without Error Bars
```{r}
repeat_cv_lasso_2 %>% 
  ggplot(aes(x = lam, y = cvm, color = seed)) + 
  geom_point() + 
  geom_line() +
  theme(legend.position = "none") +
  labs(x = expression(lambda), y = "Cross-Validation Error") +
  ggtitle("R3.6.1 glmnet 2.0-18 CV")
```


### Selected lambda
```{r}
min_cv <- repeat_cv_lasso_2 %>% 
  group_by(seed) %>% 
  filter(cvm == min(cvm))

min_cv %>% 
  ggplot(aes(y = lam, x = seed)) + 
  geom_point(aes(color = lam)) + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, size = 7)) +
  labs(x = "seed", y = expression("Cross-Validation Selected "*lambda))
```


# Group lasso

Variable grouping for group lasso

```{r}
group3 <- vector()
group3[grepl("lbx1|0",colnames(x))] <- "Non-Dioxin-like PCB"
group3[grepl("lbxd", colnames(x))] <- "TEQ"
group3[grepl("lbxf", colnames(x))] <- "TEQ"
group3[grepl("lbxh|p", colnames(x))] <- "Non-Ortho PCB"
group3[grepl("lbx118la", colnames(x))] <- "TEQ"
group3[grepl("pct", colnames(x))] <- "0"
group3[grepl("bcsi", colnames(x))] <- "0"
group3[grepl("bmi", colnames(x))] <- "0"
group3[grepl("edu", colnames(x))] <- "0"
group3[grepl("race", colnames(x))] <- "0"
group3[grepl("male", colnames(x))] <- "0"
group3[grepl("bxcot", colnames(x))] <- "0"
group3[grepl("age", colnames(x))] <- "0"
group3 <- as.factor(group3)

cbind(colnames(x), group3) #bind by columns


```

grpreg with the edited function fro grpreg that Lizzy code
```{r}
cv_lasso_3 <- cv.grpreg(x, y, group3, 
                        penalty = "grLasso", seed = 1988,
                        n.lambda = 200, max.iter = 20000)

plot(cv_lasso_3)
```


### Looping for different seeds for group lasso
```{r}
fit_with_cv = function(seed) {
  cv_lasso_3 <- cv.grpreg(x, y, group3, 
                          penalty = "grLasso", seed = seed,
                          n.lambda = 200, max.iter = 20000)
  
  #Find the lambda that results in the smallest CV error
  best_lasso_3 <- cv_lasso_3$lambda.min
  
  #Fit model with cross-validated lambda
  fit_lasso_3 <- grpreg(x, y, group3, penalty = "grLasso", lambda = best_lasso_3)
  
  fit_lasso_3
}

extract_coefs_grlasso = function(fit) {
  as_tibble(fit$beta, rownames = "variable") %>% 
    rename(beta = 2) %>% 
    filter(variable != "(Intercept)")  %>% 
    mutate(group3 = group3) %>% 
    filter(group3 %in% c("TEQ", "Non-Dioxin-like PCB", "Non-Ortho PCB"))
}

repeat_cv_group_lasso = 
  tibble(seed = 1:100) %>% 
  mutate(
    fits = map(seed, ~fit_with_cv(seed = .x)),
    coefs = map(fits, extract_coefs_grlasso)) %>% 
  dplyr::select(-fits) %>% 
  unnest() %>%
  rename(beta_grlasso = "beta")
  

```


combined data sets for plotting
```{r}
#betas <- full_join(repeat_cv_elnet, repeat_cv_lasso, by = c("variable", "seed")) %>%
 # full_join(., repeat_cv_group_lasso, by = c("variable", "seed")) 
          
```
                            
Combined plot 
```{r}
#betas_plot <- betas %>%
 # ggplot() + 
  #geom_point(aes(x = seed, y = beta_lasso, color = 'red'), size = 0.5, alpha = 0.7) + 
  #geom_point(aes(x = seed, y = beta_elnet, color = 'blue'), size = 0.5, alpha = 0.4) +
  #geom_point(aes(x = seed, y = beta_grlasso, color = 'dark green'), size = 0.5, alpha = 0.5) +
  #facet_wrap(~variable) +
  #ggtitle("Beta values for different seeds") +
  #labs(y = "beta") +
  #scale_colour_manual(name = "", 
   #     values =c('blue'='blue','red'='red', 'dark green' = 'dark green'), labels = c('Elastic Net','Lasso', 'group Lasso'))
#betas_plot
```


### beta plot for group lasso
```{r}
#inconsistend coefficients when looping the seed
repeat_cv_group_lasso %>%
  ggplot() + 
  geom_point(aes(x = seed, y = beta_grlasso), color = 'dark green', size = 0.5, alpha = 0.5) + 
  facet_wrap(~variable) +
  geom_hline(yintercept = 0) +
  ggtitle("R3.6.1 & edited grpreg 3.2-1 package")

```


