---
title: Random Seed Group Lasso Coefficients
author: "Yanelli Nunez"
date: "12/23/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

##Load packages

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse)
library(janitor)
library(caret)
library(Hmisc)
library(glmnet)
library(grpreg)
```


```{r}
sessionInfo()
```

## Data import and cleaning 

First, load the dataset; clean up names as needed; and convert factors to, well, factors. Next we remove missing values and reorder predictors (environmental variables first, confounders second). In keeping with standard practice, we'll ln-transform the environmental exposures and the outcome. This is the dataset we'll use to illustrate variable selection methods.

```{r 1}
study_pop = read_csv("./Data/studypop.csv") %>% 
  clean_names(case = c("old_janitor")) %>% 
  mutate(bmi_cat3 = as.factor(bmi_cat3),
         edu_cat = as.factor(edu_cat),
         race_cat = as.factor(race_cat),
         male = as.factor(male)) 

data = study_pop %>% 
  mutate_at(vars(contains("la")), log) %>% 
  mutate(log_telomean = log(telomean)) %>% 
  dplyr::select(log_telomean, lbx074la:lbx187la, lbxd03la:lbx194la, everything(), -seqn, -telomean) %>% 
  na.omit(log_telomean) 

# Create a matrix of predictors as x
x = model.matrix(log_telomean ~ ., data)[,-1]

# Extract outcome vector
y = data$log_telomean
```

## Variable grouping for group lasso

```{r 2}
group3 <- vector()
group3[grepl("lbx1|0",colnames(x))] <- "Non-Dioxin-like PCB"
group3[grepl("lbxd", colnames(x))] <- "TEQ"
group3[grepl("lbxf", colnames(x))] <- "TEQ"
group3[grepl("lbxh|p", colnames(x))] <- "Non-Ortho PCB"
group3[grepl("lbx118la", colnames(x))] <- "TEQ"
group3[grepl("pct", colnames(x))] <- "0"
group3[grepl("bcsi", colnames(x))] <- "0"
group3[grepl("bmi", colnames(x))] <- "0"
group3[grepl("edu", colnames(x))] <- "0"
group3[grepl("race", colnames(x))] <- "0"
group3[grepl("male", colnames(x))] <- "0"
group3[grepl("bxcot", colnames(x))] <- "0"
group3[grepl("age", colnames(x))] <- "0"
group3 <- as.factor(group3)

cbind(colnames(x), group3) #bind by columns


```

## Cross validation with a single seed

Note: we use the grpreg with the edited function from the r script file title "grpreg_warn_T.R". That file needs to be run before this codee chunk. 

```{r 3}
cv_lasso_3 <- cv.grpreg(x, y, group3, 
                        penalty = "grLasso", seed = 1988,
                        n.lambda = 200, max.iter = 20000)

plot(cv_lasso_3)
```


## Looping group lasso over 100 different seeds

Note: this code chunk takes very long to run. We saved the output data as a .csv called "summary_data_lasso.csv" that you can use in code chunk 6 and 7.

```{r 4}

# We create a function to run group lasso with differenct seeds

fit_with_cv = function(seed) {
  cv_lasso_3 <- cv.grpreg(x, y, group3, 
                          penalty = "grLasso", seed = seed,
                          n.lambda = 200, max.iter = 20000)
  
  #Find the lambda that results in the smallest CV error
  best_lasso_3 <- cv_lasso_3$lambda.min
  
  #Fit model with cross-validated lambda
  fit_lasso_3 <- grpreg(x, y, group3, penalty = "grLasso", lambda = best_lasso_3)
  
  fit_lasso_3
}

extract_coefs_grlasso = function(fit) {
  as_tibble(fit$beta, rownames = "variable") %>% 
    rename(beta = 2) %>% 
    filter(variable != "(Intercept)")  %>% 
    mutate(group3 = group3) %>% 
    filter(group3 %in% c("TEQ", "Non-Dioxin-like PCB", "Non-Ortho PCB"))
}


# We apply our function to run grlasso from 1-100 seeds

repeat_cv_group_lasso = 
  tibble(seed = 1:100) %>% 
  mutate(
    fits = map(seed, ~fit_with_cv(seed = .x)),
    coefs = map(fits, extract_coefs_grlasso)) %>% 
  dplyr::select(-fits) %>% 
  unnest() %>%
  rename(beta_grlasso = "beta")
  

```

## Tidy data obtained from running group lasso with 100 different seeds

```{r 5}

# Tidy data output from function above

data_grlasso <- repeat_cv_group_lasso %>%
  filter(!variable %in% c("bmi_cat32", "bmi_cat33", "edu_cat2", "edu_cat3", "edu_cat4", "race_cat2", "race_cat3", "race_cat4", "male1", "ln_lbxcot", "age_cent", "age_sq", "lbxwbcsi", "lbxlypct", "lbxlypct", "lbxmopct", "lbxmopct", "lbxnepct",
"lbxnepct", "lbxeopct", "lbxeopct", "lbxbapct", "lbxbapct")) %>%
  mutate(Group = ifelse(variable == "lbx118la", "mPFD", 
                        ifelse(grepl("lbx1", variable), "Non-Dioxin-like PCBs",
                          ifelse(grepl("lbx0", variable), "Non-Dioxin-like PCBs",
                            ifelse(grepl("lbxp", variable), "Non-Ortho PCBs",
                              ifelse(grepl("lbxh", variable), "Non-Ortho PCBs", "mPFD")))))) %>%
  mutate(variable = as.factor(variable),
          chem = fct_recode(variable, "PCB 74" = "lbx074la",
                                "PCB 99" = "lbx099la",
                                "PCB 118" = "lbx118la",
                                "PCB 138" = "lbx138la",
                                "PCB 153" = "lbx153la",
                                "PCB 170" = "lbx170la",
                                "PCB 180" = "lbx180la",
                                "PCB 187" = "lbx187la",
                                "PCB 194" = "lbx194la",
                                "1,2,3,6,7,8-hxcdd" = "lbxd03la",
                                "1,2,3,4,6,7,8-hpcdd" = "lbxd05la",
                               "1,2,3,4,6,7,8,9-ocdd" =  "lbxd07la",
                               "2,3,4,7,8-pncdf" =  "lbxf03la",
                               "1,2,3,4,7,8-hxcdf" =  "lbxf04la",
                               "1,2,3,6,7,8-hxcdf" =  "lbxf05la",
                               "1,2,3,4,6,7,8-hxcdf" =  "lbxf08la",
                               "PCB 169" =  "lbxhxcla",
                                "PCB 126" = "lbxpcbla")) %>%
  dplyr::select(-variable, -group3) %>%
  group_by(chem) %>%
  mutate(mean = mean(beta_grlasso),
            se = sd(beta_grlasso)/sqrt(length(beta_grlasso))) %>%
  dplyr::select(seed, Group, chem, beta_grlasso, everything())

# We count the number of non zero coefficients

counts_grlasso <- data_grlasso %>%
  filter(beta_grlasso != 0) %>%
  group_by(chem) %>%
  summarise(number_nonzero = n())
  

summary_data_grlasso <- full_join(data_grlasso, counts_grlasso, by = "chem") %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))

summary(summary_data_grlasso$beta_grlasso)

# save file for later use 
#write.csv(summary_data_grlasso, "summary_data_grlasso.csv")
```

## Visualization of beta values over different seeds (group lasso)

```{r 6 }

# load saved data
summary_data_grlasso <- read_csv("./grlasso_grpreg/summary_data_grlasso.csv") %>%
  dplyr::select(-X1)

# reorder the POPs
summary_data_grlasso$chem <- factor(summary_data_grlasso$chem,levels = c("1,2,3,6,7,8-hxcdd",
                                "1,2,3,4,6,7,8-hpcdd",
                               "1,2,3,4,6,7,8,9-ocdd",
                               "2,3,4,7,8-pncdf",
                               "1,2,3,4,7,8-hxcdf",
                               "1,2,3,6,7,8-hxcdf",
                               "1,2,3,4,6,7,8-hxcdf",
                               "PCB 118",
                               "PCB 74",
                                "PCB 99",
                                "PCB 138",
                                "PCB 153",
                                "PCB 170",
                                "PCB 180",
                                "PCB 187",
                                "PCB 194",
                                "PCB 126",
                                "PCB 169"))

# Figure
summary_data_grlasso  %>%
  ggplot() + 
  geom_point(aes(x = seed, y = beta_grlasso, color = Group), size = 0.5, alpha = 0.7) + 
  facet_wrap(~chem) +
  geom_hline(yintercept = 0) +
  theme_bw() +
labs(y = "Beta Coefficients",
       x = "Seed Number",
       title  = "Group Lasso Beta Coefficients Over 100 seeds") +
   theme(axis.text.x = element_text(size = 9.5),
                      legend.position = "bottom",
                      legend.text = element_text(size = 12),
        axis.text.y = element_text(size = 9.5),
        text = element_text(size = 12.5)) 
 


```


## Visualization summarizing the results

```{r 7}

# Figure
summary_data_grlasso %>%
  group_by(chem) %>%
  ggplot(aes(x = chem)) +
   geom_hline(yintercept = 0, color = "black") +
  geom_col(aes(y = number_nonzero/4000, fill = Group), width = 0.9, position ="identity") +
  geom_jitter(aes(y = beta_grlasso, color = Group), shape=16, size = 1.5, position=position_jitter(width = .2)) +
 # geom_col(aes(y = number_nonzero/4000, fill = Group), width = 0.8, position ="identity", alpha = 0.005) +
  scale_y_continuous(sec.axis = sec_axis(~.*4000, name = "# of Non-zero Beta Coefficients")) +
  theme_minimal(base_size = 15) +
  labs(y = "Beta Coefficients",
       x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 14),
                      legend.position = "bottom",
                      legend.text = element_text(size = 14),
        axis.text.y = element_text(size = 12),
        text = element_text(size = 12), 
        aspect.ratio = 1/1,
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_line(color = "black"),
        legend.title = element_blank(),
        axis.title.y = element_text(size = 16),
        strip.background = element_rect(fill = "white")) +
  geom_boxplot(aes(y = beta_grlasso), outlier.shape = NA, alpha = 0) +
scale_color_manual(values = c("#D55E00", "blue", "#117733")) +
   scale_fill_manual(values = c("#E69F00", "#56B4E9", "#999933"))


  
```


