---
title: Random Seed WQS weights and Index
author: "Yanelli Nunez"
date: "8/16/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth : 4
editor_options: 
  chunk_output_type: console
---

## Load packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(broom)
library(tidyverse)
library(gWQS)

options(scipen = 999)
```


```{r}
sessionInfo()
```

## Data Import and Cleaning

```{r 1}
# import the dataset
dataset = read_csv(here::here("Data/studypop.csv"))

# define the chemicals to include in the mixture
mixture = c("LBX074LA", "LBX099LA", "LBX118LA", "LBX138LA", "LBX153LA", "LBX170LA", "LBX180LA", "LBX187LA",
            "LBX194LA", "LBXD03LA", "LBXD05LA", "LBXD07LA", "LBXF03LA", "LBXF04LA", "LBXF05LA", "LBXF08LA",
            "LBXHXCLA", "LBXPCBLA")

# log-transform the outcome
dataset$log_TELOMEAN = log(dataset$TELOMEAN)
dataset$valid.var <- c(sample(c(0,1), dim(dataset)[1], replace=TRUE))



# redefine variables race_cat and edu_cat as factors
dataset$race_cat = factor(dataset$race_cat, levels = c(1, 2, 3, 4), labels = c(1, 2, 3, 4))
dataset$edu_cat = factor(dataset$edu_cat, levels = c(1, 2, 3, 4), labels = c(1, 2, 3, 4))
dataset$bmi_cat3 = factor(dataset$bmi_cat3, levels = c(1, 2, 3), labels = c(1, 2, 3))
```


## Positive adjusted model with a single seed

```{r 2}
# adjust for covariates:
# blood data: LBXWBCSI LBXLYPCT LBXMOPCT LBXEOPCT LBXBAPCT LBXNEPCT
# demographics: age_cent age_sq race_cat bmi_cat3 ln_lbxcot edu_cat


gwqs_new = gwqs(log_TELOMEAN ~ wqs + LBXWBCSI + LBXLYPCT + LBXMOPCT + LBXEOPCT + LBXBAPCT + LBXNEPCT + 
               age_cent + age_sq + race_cat + bmi_cat3 + ln_lbxcot + edu_cat + male, 
               mix_name = mixture, data = dataset, q = 10, 
               validation = 0.6, b = 5, b1_pos = TRUE, b1_constr = FALSE, 
               family = "gaussian", 
               seed = 123, plots = TRUE, tables = TRUE)

test <- tidy(gwqs_new$fit) %>% 
  filter(term == "wqs")


```


## WQS over 100 different seeds to obtain index

Note: this code chunk takes very long to run. We saved the output data as a .csv called "index_data_wqs_100seeds.csv" that you can use in code chunk 4.

```{r 3}
# Function to run wqs over 100 different seeds and extract the index optained in each seed 

fit_gwqs_index = function(seed) {
  gwqs = gwqs(log_TELOMEAN ~ wqs + LBXWBCSI + LBXLYPCT + LBXMOPCT + LBXEOPCT + LBXBAPCT + LBXNEPCT + 
               age_cent + age_sq + race_cat + bmi_cat3 + ln_lbxcot + edu_cat + male, 
               mix_name = mixture, data = dataset, q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = FALSE, 
               family = "gaussian", 
               seed = seed, plots = FALSE, tables = FALSE)
  
test <- tidy(gwqs$fit) %>% 
  filter(term == "wqs")
  
}

# Apply function with seeds 1-100 (this takes very long time to finish running. I save a .csv file with the out put data so you don't have to run this)

seed_num = 1:100
index_data  = 
  tibble(seed = seed_num) %>% 
  mutate(
    fits = map(seed, ~fit_gwqs_index(seed = .x))) %>%
  unnest() 

#save data so we don't have to run this function again because it takes too long

#write_csv(index_data, "index_data_wqs_100seeds.csv")

```


## Visualization for index across the different seeds

Note: to calculate the std.error for the pooled index we use Rubin's rule to account for both within and between seed estimates. 

```{r 4 }
index_data <- read_csv("WQS/index_data_wqs_100seeds.csv")

index_figure <- index_data %>% 
 # mutate(seed = as.character(seed)) %>%
  mutate(beta_avrg = mean(estimate),
         variance = std.error^2,
         variance_avrg = mean(variance),
         B_prep = (estimate - beta_avrg)^2,
         t = (sum(B_prep)/99) * 1.01 + variance_avrg, #calculating variance for pooled estimate using Rubin's rule
         std.error_pool = sqrt(t)) #standard error 

 
index_figure %>% 
  ggplot(aes(y = estimate, x = seed)) +
  geom_pointrange(aes(ymin = estimate - 1.96 * std.error, ymax=estimate + 1.96 * std.error), size = 0.23) + 
  geom_pointrange(aes(y = 0.0159226, x = 0,
                 ymin = 0.0159226 - 1.96 * 0.006141966, 
                 ymax = 0.0159226 + 1.96 * 0.006141966), size = 0.35, color = "red") +
 coord_flip() +
  theme_minimal(base_size = 15) + 
  xlab("Seed") +
  ylab("Estimate") +
   theme(axis.text = element_text(size = 18),
        text = element_text(size = 20))
  

```


## WQS over 100 different seeds to obtain index to obtain weights

Note: this code chunk takes very long to run. After, we tidy the data, we saved the output data as a .csv called "summary_data_wqs_100seeds.csv" that you can use in code chunks 7, 8, 9.

```{r 5}

# Function to run wqs over 100 different seeds and extract the weights optained in each seed

fit_gwqs_new = function(seed) {
  gwqs = gwqs(log_TELOMEAN ~ wqs + LBXWBCSI + LBXLYPCT + LBXMOPCT + LBXEOPCT + LBXBAPCT + LBXNEPCT + 
               age_cent + age_sq + race_cat + bmi_cat3 + ln_lbxcot + edu_cat + male, 
               mix_name = mixture, data = dataset, q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = FALSE, 
               family = "gaussian", 
               seed = seed, plots = FALSE, tables = FALSE)


weights <- as_tibble(gwqs$final_weights, rownames = "variable") %>% 
    dplyr::select(-mix_name) 

}

# Apply function with seeds 1-100 (this takes very long time to finish running. I save a .csv file with the out put data so you don't have to run this)

seed_num = 1:100


repeat_gwqs_new = 
  tibble(seed = seed_num) %>% 
  mutate(
    fits = map(seed, ~fit_gwqs_new(seed = .x))) %>%
  unnest() 


```


## Tidy data obtained from running group lasso with 100 different seeds

```{r 6}

# Tidy data 

data_gwqs <- repeat_gwqs_new%>%
  filter(!variable %in% c("bmi_cat32", "bmi_cat33", "edu_cat2", "edu_cat3", "edu_cat4", "race_cat2", "race_cat3", "race_cat4", "male1", "ln_lbxcot", "age_cent", "age_sq", "lbxwbcsi", "lbxlypct", "lbxlypct", "lbxmopct", "lbxmopct", "lbxnepct",
"lbxnepct", "lbxeopct", "lbxeopct", "lbxbapct", "lbxbapct")) %>%
  mutate(Group = ifelse(variable == "LBX118LA", "mPFD", 
                        ifelse(grepl("LBX1", variable), "Non-Dioxin-like PCBs",
                          ifelse(grepl("LBX0", variable), "Non-Dioxin-like PCBs",
                            ifelse(grepl("LBXP", variable), "Non-Ortho PCBs",
                              ifelse(grepl("LBXH", variable), "Non-Ortho PCBs", "mPFD")))))) %>%
  mutate(variable = as.factor(variable),
          chem = fct_recode(variable, "PCB 74" = "LBX074LA",
                                "PCB 99" = "LBX099LA",
                                "PCB 118" = "LBX118LA",
                                "PCB 138" = "LBX138LA",
                                "PCB 153" = "LBX153LA",
                                "PCB 170" = "LBX170LA",
                                "PCB 180" = "LBX180LA",
                                "PCB 187" = "LBX187LA",
                                "PCB 194" = "LBX194LA",
                                "1,2,3,6,7,8-hxcdd" = "LBXD03LA",
                                "1,2,3,4,6,7,8-hpcdd" = "LBXD05LA",
                               "1,2,3,4,6,7,8,9-ocdd" =  "LBXD07LA",
                               "2,3,4,7,8-pncdf" =  "LBXF03LA",
                               "1,2,3,4,7,8-hxcdf" =  "LBXF04LA",
                               "1,2,3,6,7,8-hxcdf" =  "LBXF05LA",
                               "1,2,3,4,6,7,8-hxcdf" =  "LBXF08LA",
                               "PCB 169" =  "LBXHXCLA",
                                "PCB 126" = "LBXPCBLA")) %>%
  dplyr::select(-variable) %>%
  group_by(chem) %>%
  mutate(total_mean = mean(mean_weight),
            sd_total_mean = sd(mean_weight)) %>%
  dplyr::select(seed, Group, chem, mean_weight, everything())

summary(data_gwqs)

# We count the number of non zero coefficients

counts <- data_gwqs %>%
  filter(mean_weight >= 0.06) %>% # 1/18 = 0.06 (1 divided by the number of chemicals)
  group_by(chem) %>%
  summarise(above_threshold = n())
  
summary_data_gwqs <- full_join(data_gwqs, counts, by = "chem") %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))


#write_csv(summary_data_gwqs, "summary_data_wqs_100seeds.csv")

```



## Weights plots for WQS

```{r 7 }
summary_data_gwqs = read_csv("./WQS/summary_data_wqs_100seeds.csv")

summary_data_gwqs %>%
  ggplot() + 
  geom_point(aes(x = seed, y = mean_weight), color = "red", size = 0.5, alpha = 0.5) + 
  facet_wrap(~chem) + 
  geom_hline(yintercept = 0) +
  ggtitle("Chem weights over 100 different seeds") 

```

## Visualization of beta values over different seeds 

fig size: 7 x 8.5 in
```{r 8}

summary_data_gwqs$chem <- factor(summary_data_gwqs$chem,levels = c("1,2,3,6,7,8-hxcdd",
                                "1,2,3,4,6,7,8-hpcdd",
                               "1,2,3,4,6,7,8,9-ocdd",
                               "2,3,4,7,8-pncdf",
                               "1,2,3,4,7,8-hxcdf",
                               "1,2,3,6,7,8-hxcdf",
                               "1,2,3,4,6,7,8-hxcdf",
                               "PCB 118",
                               "PCB 74",
                                "PCB 99",
                                "PCB 138",
                                "PCB 153",
                                "PCB 170",
                                "PCB 180",
                                "PCB 187",
                                "PCB 194",
                                "PCB 126",
                                "PCB 169"))
summary_data_gwqs %>%
  group_by(chem) %>%
  ggplot(aes(x = chem)) +
  geom_col(aes(y = above_threshold/150, fill = Group), width = 0.9, position ="identity") +
  geom_jitter(aes(y = mean_weight, color = Group), shape=16, size = 1.5, position=position_jitter(width = .2)) +
   geom_hline(yintercept = 0.06, size = 0.7) +
  scale_y_continuous(sec.axis = sec_axis(~.*150, name = "# of Repeats Above Threshold")) +
  theme_minimal(base_size = 15) +
  labs(y = "Weight",
       x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 14),
                      legend.position = "bottom",
                      legend.text = element_text(size = 14),
        axis.text.y = element_text(size = 12),
        text = element_text(size = 12), 
        aspect.ratio = 1/1,
        axis.line.x = element_blank(),
       # axis.line.y = element_line(color = "gray"),
        axis.ticks.x = element_blank(),
       legend.title = element_blank(),
        axis.title.y = element_text(size = 16)) +
        # strip.background = element_rect(fill = "white")) +
  geom_boxplot(aes(y = mean_weight), outlier.shape = NA, alpha = 0) +
  scale_color_manual(values = c("#D55E00", "blue", "#117733")) +
   scale_fill_manual(values = c("#E69F00", "#56B4E9", "#999933")) 
 
 
```

## Largest weight values 

```{r 9}
max_weight <- summary_data_gwqs %>%
  group_by(seed) %>%
  filter(mean_weight == max(mean_weight)) %>%
  group_by(chem) %>%
  summarise(n())

max_weight

```






