---
title: "betas for WQS based on seed number"
author: "Yanelli Nunez"
date: "8/16/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth : 4
editor_options: 
  chunk_output_type: console
---

#Questions for Marianthi:

Lasso & group lasso use the random number genrador for the cross validation part to chose the best lambda value. Lambda can be any number from 0 -> infinity. The random number generator, produces a grid of numbers that are then tested to identify the lambda that results in the smallest prediction error. 
-Where does WQS needs a seed? does it use a tunning parameter? is this in the Lagran multiplier part? 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
#install.packages("gWQS", dependencies = TRUE)
#devtools::package_deps("gWQS") can use this to see which package dependencies you already have

library(tidyverse)

#install.packages("devtools")
#require(devtools)
#install_version("gWQS", version = "1.1.1", repos = "http://cran.us.r-project.org")

library(gWQS)

options(scipen = 999)
```


```{r}
sessionInfo()
```

## Data Import and Cleaning

```{r import}
# import the dataset
dataset = read_csv(here::here("Data/studypop.csv"))

# define the chemicals to include in the mixture
mixture = c("LBX074LA", "LBX099LA", "LBX118LA", "LBX138LA", "LBX153LA", "LBX170LA", "LBX180LA", "LBX187LA",
            "LBX194LA", "LBXD03LA", "LBXD05LA", "LBXD07LA", "LBXF03LA", "LBXF04LA", "LBXF05LA", "LBXF08LA",
            "LBXHXCLA", "LBXPCBLA")

# log-transform the outcome
dataset$log_TELOMEAN = log(dataset$TELOMEAN)
dataset$valid.var <- c(sample(c(0,1), dim(dataset)[1], replace=TRUE))



# redefine variables race_cat and edu_cat as factors
dataset$race_cat = factor(dataset$race_cat, levels = c(1, 2, 3, 4), labels = c(1, 2, 3, 4))
dataset$edu_cat = factor(dataset$edu_cat, levels = c(1, 2, 3, 4), labels = c(1, 2, 3, 4))
dataset$bmi_cat3 = factor(dataset$bmi_cat3, levels = c(1, 2, 3), labels = c(1, 2, 3))
```

## Weighted Quantile Sums
Note:
There is a new version of the gwqs package (version 2.0.0). The input for the function is a bit different compare to version 1.1.1 and the results are different. Here I am using the old package to check the seed issue in the first code chunk and the new package in the following code chunk.

For both packages the solnp optimatimation function is use to estimate the weights at each bootstrop


### Positive Adjusted Model
### For old gWQS version

Here we obtained the same results that we optain in the workshop using the old package with the new version of R
```{r code for old package}
# adjust for covariates:
# blood data: LBXWBCSI LBXLYPCT LBXMOPCT LBXEOPCT LBXBAPCT LBXNEPCT
# demographics: age_cent age_sq race_cat bmi_cat3 ln_lbxcot edu_cat


result_old = gwqs(log_TELOMEAN ~ LBXWBCSI + LBXLYPCT + LBXMOPCT + LBXEOPCT + LBXBAPCT + LBXNEPCT + 
               age_cent + age_sq + race_cat + bmi_cat3 + ln_lbxcot + edu_cat + male, 
               mix_name = mixture, data = dataset, q = 10, 
               validation = 0.6, valid_var = NULL, b = 100, b1_pos = TRUE, b1_constr = FALSE, 
               family = "gaussian", 
               seed = 123, wqs2 = FALSE, plots = TRUE, tables = TRUE)

summary(result_old$fit)
result_old$final_weights %>% knitr::kable()
confint(result_old$fit)

```


```{r function code for old package}

fit_gwqs = function(seed) {
  gwqs = gwqs(log_TELOMEAN ~ LBXWBCSI + LBXLYPCT + LBXMOPCT + LBXEOPCT + LBXBAPCT + LBXNEPCT + 
               age_cent + age_sq + race_cat + bmi_cat3 + ln_lbxcot + edu_cat + male, 
               mix_name = mixture, data = dataset, q = 10, 
               validation = 0.6, valid_var = NULL, b = 100, b1_pos = TRUE, b1_constr = FALSE, 
               family = "gaussian", 
               seed = seed, wqs2 = FALSE, plots = FALSE, tables = FALSE)


weights <- as_tibble(gwqs$final_weights, rownames = "variable") %>% 
    select(-mix_name) 
  weights

}


repeat_gwqs = 
  tibble(seed = 1:50) %>% 
  mutate(
    fits = map(seed, ~fit_gwqs(seed = .x))) %>%
  unnest() 

write_csv(repeat_gwqs, "gWQS_1.1.1_R3.6.1 weights.csv")
```


```{r}
repeat_gwqs %>%
  ggplot() + 
  geom_point(aes(x = seed, y = mean_weight), color = "red", size = 0.5, alpha = 0.5) + 
  facet_wrap(~variable) + 
  geom_hline(yintercept = 0) +
  ggtitle("R3.6.1 & gWQS 1.1.1")
```


### Positive Adjusted Model
### for new gWQS version

```{r code for new package}
# adjust for covariates:
# blood data: LBXWBCSI LBXLYPCT LBXMOPCT LBXEOPCT LBXBAPCT LBXNEPCT
# demographics: age_cent age_sq race_cat bmi_cat3 ln_lbxcot edu_cat


gwqs_new = gwqs(log_TELOMEAN ~ wqs + LBXWBCSI + LBXLYPCT + LBXMOPCT + LBXEOPCT + LBXBAPCT + LBXNEPCT + 
               age_cent + age_sq + race_cat + bmi_cat3 + ln_lbxcot + edu_cat + male, 
               mix_name = mixture, data = dataset, q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = FALSE, 
               family = "gaussian", 
               seed = 123, plots = TRUE, tables = TRUE)


summary(gwqs_new$fit)
gwqs_new$final_weights %>% knitr::kable()
```


```{r function code for new package, cache = TRUE}

fit_gwqs_new = function(seed) {
  gwqs = gwqs(log_TELOMEAN ~ wqs + LBXWBCSI + LBXLYPCT + LBXMOPCT + LBXEOPCT + LBXBAPCT + LBXNEPCT + 
               age_cent + age_sq + race_cat + bmi_cat3 + ln_lbxcot + edu_cat + male, 
               mix_name = mixture, data = dataset, q = 10, 
               validation = 0.6, b = 100, b1_pos = TRUE, b1_constr = FALSE, 
               family = "gaussian", 
               seed = seed, plots = FALSE, tables = FALSE)


weights <- as_tibble(gwqs$final_weights, rownames = "variable") %>% 
    dplyr::select(-mix_name) 
  weights

}

seed_num = 1:100 



repeat_gwqs_new = 
  tibble(seed = seed_num) %>% 
  mutate(
    fits = map(seed, ~fit_gwqs_new(seed = .x))) %>%
  unnest() 


```


#Organizing data 

```{r}
data_gwqs <- repeat_gwqs_new%>%
  filter(!variable %in% c("bmi_cat32", "bmi_cat33", "edu_cat2", "edu_cat3", "edu_cat4", "race_cat2", "race_cat3", "race_cat4", "male1", "ln_lbxcot", "age_cent", "age_sq", "lbxwbcsi", "lbxlypct", "lbxlypct", "lbxmopct", "lbxmopct", "lbxnepct",
"lbxnepct", "lbxeopct", "lbxeopct", "lbxbapct", "lbxbapct")) %>%
  mutate(Group = ifelse(variable == "LBX118LA", "mPFD", 
                        ifelse(grepl("LBX1", variable), "Non-Dioxin-like PCBs",
                          ifelse(grepl("LBX0", variable), "Non-Dioxin-like PCBs",
                            ifelse(grepl("LBXP", variable), "Non-Ortho PCBs",
                              ifelse(grepl("LBXH", variable), "Non-Ortho PCBs", "mPFD")))))) %>%
  mutate(variable = as.factor(variable),
          chem = fct_recode(variable, "PCB 74" = "LBX074LA",
                                "PCB 99" = "LBX099LA",
                                "PCB 118" = "LBX118LA",
                                "PCB 138" = "LBX138LA",
                                "PCB 153" = "LBX153LA",
                                "PCB 170" = "LBX170LA",
                                "PCB 180" = "LBX180LA",
                                "PCB 187" = "LBX187LA",
                                "PCB 194" = "LBX194LA",
                                "1,2,3,6,7,8-hxcdd" = "LBXD03LA",
                                "1,2,3,4,6,7,8-hpcdd" = "LBXD05LA",
                               "1,2,3,4,6,7,8,9-ocdd" =  "LBXD07LA",
                               "2,3,4,7,8-pncdf" =  "LBXF03LA",
                               "1,2,3,4,7,8-hxcdf" =  "LBXF04LA",
                               "1,2,3,6,7,8-hxcdf" =  "LBXF05LA",
                               "1,2,3,4,6,7,8-hxcdf" =  "LBXF08LA",
                               "PCB 169" =  "LBXHXCLA",
                                "PCB 126" = "LBXPCBLA")) %>%
  dplyr::select(-variable) %>%
  group_by(chem) %>%
  mutate(total_mean = mean(mean_weight),
            sd_total_mean = sd(mean_weight)) %>%
  dplyr::select(seed, Group, chem, mean_weight, everything())

summary(data_gwqs)

#counting non zero coefficients

counts <- data_gwqs %>%
  filter(mean_weight >= 0.06) %>% # 1/18 = 0.06 (1 divided by the number of chemicals)
  group_by(chem) %>%
  summarise(above_threshold = n())
  
summary_data_gwqs <- full_join(data_gwqs, counts, by = "chem") %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))


write_csv(summary_data_gwqs, "summary_data_wqs_100seeds.csv")

```


#Weights plots for WQS
```{r}

repeat_gwqs_new %>%
  ggplot() + 
  geom_point(aes(x = seed, y = mean_weight), color = "red", size = 0.5, alpha = 0.5) + 
  facet_wrap(~variable) + 
  geom_hline(yintercept = 0) +
  ggtitle("R3.6.1 & gWQS 2.0") 
 # theme(axis.text.x = element_text(angle = 90, size = 4))
```




#final summary plot for lasso
```{r}

summary_data_gwqs = read_csv("./WQS/summary_data_wqs_100seeds.csv")
summary_data_gwqs$chem <- factor(summary_data_gwqs$chem,levels = c("1,2,3,6,7,8-hxcdd",
                                "1,2,3,4,6,7,8-hpcdd",
                               "1,2,3,4,6,7,8,9-ocdd",
                               "2,3,4,7,8-pncdf",
                               "1,2,3,4,7,8-hxcdf",
                               "1,2,3,6,7,8-hxcdf",
                               "1,2,3,4,6,7,8-hxcdf",
                               "PCB 118",
                               "PCB 74",
                                "PCB 99",
                                "PCB 138",
                                "PCB 153",
                                "PCB 170",
                                "PCB 180",
                                "PCB 187",
                                "PCB 194",
                                "PCB 126",
                                "PCB 169"))
summary_data_gwqs %>%
  group_by(chem) %>%
  ggplot(aes(x = chem)) +
  geom_hline(yintercept = 0.06, size = 0.7) +
  geom_jitter(aes(y = mean_weight, color = Group), shape=16, size = 1, alpha = 1,  position=position_jitter(width = .2)) +
  #geom_point(aes(y = total_mean, fill = Group), shape = 23, size = 3, alpha = 1) +
  #geom_errorbar(aes(ymin = total_mean-sd_total_mean, ymax=total_mean+sd_total_mean, group = chem), width=0.1, size = 0.25) +
 #stat_summary(aes(y = mean_weight), fun.data = mean_sdl, fun.args = list(mult = 1), geom = "errorbar", width=0.25) +
  #stat_summary(aes(y = mean_weight, fill = Group), fun.y = mean, geom ="point", shape=23, size=3, alpha = 1) +
  geom_col(aes(y = above_threshold/150, fill = Group), width = 0.8, position ="identity", alpha = 0.002) +
  scale_y_continuous(sec.axis = sec_axis(~.*150, name = "# of Repeats Above Threshold")) +
  theme_classic() +
  labs(y = "Weight",
       x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 9.5),
                      legend.position = "bottom",
                      legend.text = element_text(size = 12),
        axis.text.y = element_text(size = 9.5),
        text = element_text(size = 12.5), 
        aspect.ratio = 1/1,
        axis.line.x = element_blank(),
       # axis.line.y = element_line(color = "gray"),
        axis.ticks.x = element_blank()) +
  geom_boxplot(aes(y = mean_weight), outlier.shape = NA, alpha = 0) 
  
 
 
```


#Specifying the splits for the training and validation sets
```{fixed splits, cache = TRUE}

fit_gwqs_new_split = function(seed) {
  gwqs = gwqs(log_TELOMEAN ~ wqs + LBXWBCSI + LBXLYPCT + LBXMOPCT + LBXEOPCT + LBXBAPCT + LBXNEPCT + 
               age_cent + age_sq + race_cat + bmi_cat3 + ln_lbxcot + edu_cat + male, 
               mix_name = mixture, data = dataset, q = 10, 
               valid_var = "valid.var", b = 100, b1_pos = TRUE, b1_constr = FALSE, 
               family = "gaussian", 
               seed = seed, plots = FALSE, tables = FALSE)


weights <- as_tibble(gwqs$final_weights, rownames = "variable") %>% 
    select(-mix_name) 
  weights

}


#run function

seed_num = 1:50 

repeat_gwqs_new_split = 
  tibble(seed = seed_num) %>% 
  mutate(
    fits = map(seed, ~fit_gwqs_new_split(seed = .x))) %>%
  unnest() 

#plot
repeat_gwqs_new_split %>%
  ggplot() + 
  geom_point(aes(x = seed, y = mean_weight), color = "red", size = 0.5, alpha = 0.5) + 
  facet_wrap(~variable) + 
  geom_hline(yintercept = 0) +
  ggtitle("R3.6.1 & gWQS 2.0 (fixed test/validation group)") +
  ylab("weights")


```





